\bf Comparing the Power of Plot Designs to Reveal Correlation

### 2015-12-10 paper/paper.Rnw
got it to compile, almost its complaining about latex commands now, will clean this up on plane

\documentclass[12pt]{article}
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[colorlinks=TRUE, linkcolor=blue]{hyperref}
\usepackage{wrapfig,float}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{ulem}
\usepackage[section]{placeins}
\usepackage{afterpage}

\graphicspath{{images/}}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

@@

\newpage
\spacingset{1.45} % DON'T change the spacing!
\newpage
<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(tidyr)
# Convert reshape2 to tidyr, melt to gather - done
#library(reshape2)
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
#library(splitstackshape) - changed cSplit to unnest
library(readr)
library(lubridate)
# Convert these to read_csv at some point
correl <- read.csv("correlations.csv")
full <- read.csv("full.csv", row.names=1) 
Turk <- read.csv("turk.csv")

@@

\begin{figure}[htbp]
  \centering
  \includegraphics{cny line}
  \caption{Exchange rates for Australian dollars and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid time series. (Both sets of values were standardized to bring them to the same scale for plotting.) What would you report the correlation between the two series to be?}
  \label{cny}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}
    \includegraphics{nzd scatter}
  \end{subfigure}
  \begin{subfigure}
    \includegraphics{nzd line}
  \end{subfigure}
  \caption{ Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right). Which one can you read the correlation from more accurately? Point-wise correlation between the two series, ignoring autodependence, is 0.64.}
  \label{nzd}
\end{figure}

To decide on an appropriate display requires awareness of cognitive principles in psychology such as the Gestalt law of common fate described by Wertheimer in 1923 [22, 26]. This law describes a tendency in human perception to 'see' a positive relationship between two objects that briefly move together over time, leading people to believe that they stay linked and share a 'common fate' (Figure~\ref{cny}), even if no relationship exists. Lee and Blake's [13] work finds that humans are able to perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that a graphic than can utilise a dimension for time may be able to use the information provided by temporal dependence and potentially outperform the graph types that do not, such as the scatter plot [18]. However, keep in mind magical thinking [6], the desire to find any pattern may lead to a false discovery of positive correlation. The law then has two effects, if positive correlation exists in temporal data, the observer may be drawn to it easier. If positive correlation does not exist, the observer may be fooled into thinking it does anyway. Either effect would lead to a more confident rejection of the null than if the same data was displayed as a scatter plot. Figure~\ref{cny} highlights the dangerous interaction between 'magical thinking' and the Gestalt Law, the observer is drawn to the period from April 8 to April 22, and may falsely believe that the two currencies may continue to move together. However, over the entire series the linear correlation between the two currencies is -0.12, indicating there is no actual relationship. Incorrectly relying on two currencies, or many other sets of economic variables, being dependent on each other can be extremely dangerous for a firm. Work by Robbins [19] is similarly critical of the time-based line graph, where it is argued that information from overlaid line graphs can easily be misleading (Figure~\ref{murder}). When both lines dive after 1993, it is natural to compare the horizontal distance and say that the difference between the lines in 1994 is small. However, this is incorrect, as looking at the small horizontal distance is comparing the lines at different dates. It is not obvious that the difference in 1994 is the second greatest difference across the entire range, only slightly smaller than in 1993. Vanderplas [25] examines this effect in more detail. It is dubbed the 'Sine Illusion', where the human eye is poor at interpreting lines with such a sharp slope. As these problems are unique to the line graph, they argue that the scatter plot, which does not have any ability to display time, is the superior choice. However, in practice many data analysts are split between the two major alternatives for temporal data. The overlaid line graph is justified by its ability to present more information to the analyst; but many argue that it is this extra information that is misleading, and revert to the scatter plot for its strong non-temporal performance.

\begin{figure}[htbp]
  \centering
  \includegraphics{murders}
  \caption{[19] Murders in New York State (Orange) and New York City (Blue). The difference between the two represents murders outside of the city. What size would you judge the difference to be in 1994?}
  \label{murder}
\end{figure}

@@

If all $K$ observers pick the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ . This research used $m= 20$, giving us a significance level (and Type 1 error rate) of $α = 0.05$. The plot of 'true' data is placed randomly amongst 19 null plots to form the lineup.
\begin{figure}[htbp]
  \centering
<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(1234)
a <- gen_true_data(24, 0.95)
td <- gather(a, variable, value, c(X1, X2))
b <- gen_null(24)
b$.n <- NULL
nd_l <- gather(b, variable, value, c(X1, X2))
nd_l$.n <- rep(1:19, rep(24, 19))
pos <- 7
td <- data.frame(.n=rep(pos, nrow(td)), td)
lg <- nd_l$.n < pos
nd_l1 <- data.frame(.n=nd_l$.n[lg], t=nd_l$t[lg], variable=nd_l$variable[lg], 
                    value=nd_l$value[lg])
nd_l2 <- data.frame(.n=nd_l$.n[!lg], t=nd_l$t[!lg], variable=nd_l$variable[!lg], 
                    value=nd_l$value[!lg])
nd_l2$.n <- nd_l2$.n + 1
d <- rbind(nd_l1, td, nd_l2)

@@

\begin{figure}[htbp]
  \centering
<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(123)
uns <- gen_true_data(96, 0.8)
uns_l <- gather(uns, variable, value, -t)
p1a <- ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() +labs(title="Unsmoothed Overlaid Lines") +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3a <- ggplot(data=uns, aes(x=X1, y=X2)) + geom_point()+ labs(title="Unsmoothed Scatterplot)") +
    theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")

set.seed(123)
sm <- gen_true_data(96, 0.8, smoothed = TRUE)
sm_l <- gather(sm, variable, value, -t)
p2a <- ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Smoothed Overlaid Lines") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4a <- ggplot(data=sm, aes(x=X1, y=X2)) + geom_point() + labs(title="Smoothed Scatterplot") +
    theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")

@@

\begin{figure}[htbp]
  \centering
<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(101)
a <- gen_true_data(24, 0.95)
b <- gen_true_data(24, -0.95)
ca <- cor(a$X1, a$X2)
cb <- cor(b$X1, b$X2)
a_l <- gather(a, variable, value, -t)
b_l <- gather(b, variable, value, -t)
p1 <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p2 <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3 <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4 <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
grid.arrange(p1, p2, p3, p4, ncol=2)
@
  \label{designs}
  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
\end{figure}

@@

Figure~\ref{results} shows the raw results of the survey and Figure~\ref{powerdiff} shows the difference in power with a $95\%$ confidence interval calculated according to the formula in \ref{eq:2}. Values above zero indicate that the scatter plot had more power to detect the true plot than the overlaid lines.
Performance between the two plot designs appears to be similar for positive correlation, with the scatter plot power gradually improving relative to the overlaid lines as sample size increases. The scatter plot appears to be the superior choice as it retains its power with negative correlation, where the overlaid lines performance suffered. This is not surprising, as positive and negative correlation appear as symmetric graphs in the scatter plot but the two characteristics appear very differently in the overlaid line graph (Figure~\ref{designs}). The right side shows the performance benefits of smoothing overlaid line time series data. Much of the 'noise' in the time series is removed and studying the overall trend shows a greatly improved ability to successfully detect the true data plot when it is negatively correlated. The scatter plot still appears to be the superior plot, but much of the gap between the two is reduced for positive correlation.

@@

Figure~\ref{samplesize} demonstrates the effects of changing sample size for unsmoothed data. The scatter plot has increased power to detect correlation almost universally for any increase. Sample size influences the overlaid lines differently, they appears to benefit for positively correlated true plots, but it does not improve the weak performance in negative correlation.
Figure~\ref{time} and Figure~\ref{confidence} give a breakdown of the interactions between the response variables: plot detection, time and confidence. Figure 8 is the histogram of the time taken to record a response, split by plot design and sign of correlation and true plot detection. The time data was highly positively skewed with a maximum of 1760 seconds so the variable was log-transformed. We found that the subject could identify the true plot out of the overlaid line graph 33.6\% of the time and it took on average $e^{3.442}$ = 31.2 seconds to make a choice. Successful detections were slightly faster than unsuccessful decision, taking on average 28.7 seconds to pick the true plot and 32.6 seconds if a null was selected. The scatter plots performed much better, with the true plot picked 52.9\% of the time, with an average of 22.7 seconds required to make a decision. If the true plot was detected it took only 18.6 seconds, and 29.3 seconds for a null plot selection.

@@

Next we analysed the individual ability of each of our subjects. Figure~\ref{confidence2} plots the accuracy of each subject on the y axis versus the average confidence on the x axis. While we expected that subjects with greater visual skills would report lower values of confidence, there does not appear to be any relationship between the two. The accuracy is normally distributed (Jarque-Bera p value = 0.418), with a mean of 42.5\% and a standard deviation of 17.8\%. The strongest subject successfully identified the true plot 91\% of the time, whilst the weakest was never able to identify the true plot. As all subjects were shown a balanced mix of true plot correlations, the extreme difference in individual visual ability is striking. Visual ability will be controlled via a subject specific random effect in the models in the next section.

@@

There are two major reasons why a subject may fail to detect the true plot out of a lineup. Either the plot design was unable to display correlation for that particular combination of factors (and the analyst would commit a Type II error by not rejecting the null hypothesis), or a null plot that had a greater correlation than the true plot was selected instead (and the analyst would commit a Type I error by rejecting the null hypothesis). We next can analyse these kinds of errors with the pin plots found in Figure~\ref{pins}. The frequency of selection of each plot in each lineup, either true or null, is represented by a pin, placed at the value of the correlation between the two variables shown in that plot. The correlation of the selected plot is on the x-axis and the frequency of selection is plotted on the y axis. The correlation of the simulated (true) data (r=-0.9, -0.7, -0.3, 0.3, 0.7, 0.9) in the plot, the sample size (t=12, 24, 48, 96), and the plot design (unsmoothed/smoothed, scatterplot/lineplot) is used to facet the plots. The blue pins indicate the true plots and the red pins indicate the null plots. If the plot design has the power to display correlation, we would expect the highest correlation plot to be selected, leading to a Type I error where a plot from the null distribution spuriously had high correlation. We find that this is generally not the case. The selections appear to have a wide range of correlations, indicating that if a subject selected a null plot it was typically not due to excess spurious correlation. Instead the failure to select the highest correlated plot indicates that a Type II error was committed rather than Type I as the plot designs did not have the power to differentiate between medium and low correlation plots. However, for the scatter plot, whenever there was a plot with a very strong correlation, usually in the form of the true plot, it was overwhelmingly selected. For the overlaid lines this mostly occurred only for plots with strong positive correlation, which is consistent with its general lack of power to reveal negative relationships in data.

\begin{figure}[htbp]
  \centering
<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
# correl <- read.csv("correlations.csv") # You have already read this
pin <- select(full, pic_id, response_0, actual_location, 
              test_param, smoothed, n) #full[,c(5, 6, 10, 15, 16, 17)]
# Need to use the tidyr commands instead of cSplit
# Here's where we will need to make some changes to incorporate the multiple picks
pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
counts <- data.frame(table(pin$pic_id, pin$response_0))
counts$Var1 <- as.integer(counts$Var1)
counts$Var2 <- as.integer(counts$Var2)
colnames(counts) <- c("lineup", "plot", "count")
pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
pin$r <- factor(pin$r)
#colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
pin <- mutate(pin, design=ifelse(test_param=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
pin$base <- 0
pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))
pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
pin <- mutate(pin, percent = count/attempt)
ggplot(data=subset(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")

@@

\begin{figure}[htbp]
  \centering
  \includegraphics{anom1}
  \label{anom1}
  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707. The true plot in position 8 was not detected in five evaluations. However, plot 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225. When asked for a reason of their selections, subjects answered that the lines were closest together.}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics{anom2}
  \label{anom2}
  \caption{A lineup of unsmoothed scatter plots with a sample size of 24. The true plot is located in position 4 with a correlation of -0.491, and was selected by one out of eleven observers (p = 0.33, see formula \ref{eq:1}). The plot in position 6 was selected by nine of these eleven observers (p < 0.001) despite having a correlation of -0.037. This null plot shows a unique clustering effect that draws in the observer despite not having any linear relationship in the data, an example of the over-interpretation of randomness. Subjects most often cited the clustering pattern formed as the reason for their selection. The final selection was position 8, with a correlation of -0.226 (p = 0.33).}
\end{figure}

>>>>

\documentclass[12pt]{article}
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[colorlinks=TRUE, linkcolor=blue]{hyperref}
\usepackage{wrapfig,float}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{amsmath}
%\usepackage{dsfont}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{ulem}
%\usepackage[section]{placeins}
\usepackage{afterpage}

\graphicspath{{figure/}}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

@@

\newpage
\spacingset{1.45} % DON'T change the spacing!
\newpage

<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
opts_chunk$set(cache=FALSE)
library(lme4)
library(ggplot2)
library(mnormt)
library(reshape2)
# Convert reshape2 to tidyr, melt to gather
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
library(splitstackshape)
library(readr)
library(lubridate)
# Convert these to read_csv at some point
correl <- read.csv("correlations.csv")
full <- read.csv("full.csv", row.names=1) 
Turk <- read.csv("turk.csv")

@@

Insert Figure 1

@@

Insert Figure 2

To decide on an appropriate display requires awareness of cognitive principles in psychology such as the Gestalt law of common fate described by Wertheimer in 1923 [22, 26]. This law describes a tendency in human perception to 'see' a positive relationship between two objects that briefly move together over time, leading people to believe that they stay linked and share a 'common fate' (Figure 1), even if no relationship exists. Lee and Blake's [13] work finds that humans are able to perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that a graphic than can utilise a dimension for time may be able to use the information provided by temporal dependence and potentially outperform the graph types that do not, such as the scatter plot [18]. However, keep in mind magical thinking [6], the desire to find any pattern may lead to a false discovery of positive correlation. The law then has two effects, if positive correlation exists in temporal data, the observer may be drawn to it easier. If positive correlation does not exist, the observer may be fooled into thinking it does anyway. Either effect would lead to a more confident rejection of the null than if the same data was displayed as a scatter plot. Figure 1 highlights the dangerous interaction between 'magical thinking' and the Gestalt Law, the observer is drawn to the period from April 8 to April 22, and may falsely believe that the two currencies may continue to move together. However, over the entire series the linear correlation between the two currencies is -0.12, indicating there is no actual relationship. Incorrectly relying on two currencies, or many other sets of economic variables, being dependent on each other can be extremely dangerous for a firm. Work by Robbins [19] is similarly critical of the time-based line graph, where it is argued that information from overlaid line graphs can easily be misleading (Figure 3). When both lines dive after 1993, it is natural to compare the horizontal distance and say that the difference between the lines in 1994 is small. However, this is incorrect, as looking at the small horizontal distance is comparing the lines at different dates. It is not obvious that the difference in 1994 is the second greatest difference across the entire range, only slightly smaller than in 1993. Vanderplas [25] examines this effect in more detail. It is dubbed the 'Sine Illusion', where the human eye is poor at interpreting lines with such a sharp slope. As these problems are unique to the line graph, they argue that the scatter plot, which does not have any ability to display time, is the superior choice. However, in practice many data analysts are split between the two major alternatives for temporal data. The overlaid line graph is justified by its ability to present more information to the analyst; but many argue that it is this extra information that is misleading, and revert to the scatter plot for its strong non-temporal performance.

Insert Figure 3

@@

If all $K$ observers pick the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ . This research used $m= 20$, giving us a significance level (and Type 1 error rate) of $α = 0.05$. The plot of 'true' data is placed randomly amongst 19 null plots to form the lineup.
\begin{figure}[htbp]
  \centering
<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(1234)
a <- gen_true_data(24, 0.95)
td <- melt(a, id="t") 
b <- gen_null(24)
b$.n <- NULL
nd_l <- melt(b, id="t")
nd_l$.n <- rep(1:19, rep(24, 19))
pos <- 7
td <- data.frame(.n=rep(pos, nrow(td)), td)
lg <- nd_l$.n < pos
nd_l1 <- data.frame(.n=nd_l$.n[lg], t=nd_l$t[lg], variable=nd_l$variable[lg], 
                    value=nd_l$value[lg])
nd_l2 <- data.frame(.n=nd_l$.n[!lg], t=nd_l$t[!lg], variable=nd_l$variable[!lg], 
                    value=nd_l$value[!lg])
nd_l2$.n <- nd_l2$.n + 1
d <- rbind(nd_l1, td, nd_l2)

@@

\begin{figure}[htbp]
  \centering
<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(123)
uns <- gen_true_data(96, 0.8)
uns_l <- melt(uns, id="t")
p1a <- ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() +labs(title="Unsmoothed Overlaid Lines") +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3a <- ggplot(data=uns, aes(x=X1, y=X2)) + geom_point()+ labs(title="Unsmoothed Scatterplot)") +
    theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")

set.seed(123)
sm <- gen_true_data(96, 0.8, smoothed = TRUE)
sm_l <- melt(sm, id="t")
p2a <- ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Smoothed Overlaid Lines") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4a <- ggplot(data=sm, aes(x=X1, y=X2)) + geom_point() + labs(title="Smoothed Scatterplot") +
    theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")

@@

\begin{figure}[htbp]
  \centering
<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
set.seed(101)
a <- gen_true_data(24, 0.95)
b <- gen_true_data(24, -0.95)
ca <- cor(a$X1, a$X2)
cb <- cor(b$X1, b$X2)
a_l <- melt(a, id="t")
b_l <- melt(b, id="t")
p1 <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p2 <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3 <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4 <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
grid.arrange(p1, p2, p3, p4, ncol=2)
@
  \label{designs}
  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
\end{figure}

@@

Figure 7 shows the raw results of the survey and Figure 8 shows the difference in power with a $95\%$ confidence interval calculated according to the formula in \ref{eq:2}. Values above zero indicate that the scatter plot had more power to detect the true plot than the overlaid lines.
Performance between the two plot designs appears to be similar for positive correlation, with the scatter plot power gradually improving relative to the overlaid lines as sample size increases. The scatter plot appears to be the superior choice as it retains its power with negative correlation, where the overlaid lines performance suffered. This is not surprising, as positive and negative correlation appear as symmetric graphs in the scatter plot but the two characteristics appear very differently in the overlaid line graph (Figure 6). The right side shows the performance benefits of smoothing overlaid line time series data. Much of the 'noise' in the time series is removed and studying the overall trend shows a greatly improved ability to successfully detect the true data plot when it is negatively correlated. The scatter plot still appears to be the superior plot, but much of the gap between the two is reduced for positive correlation.

@@

Figure 9 demonstrates the effects of changing sample size for unsmoothed data. The scatter plot has increased power to detect correlation almost universally for any increase. Sample size influences the overlaid lines differently, they appears to benefit for positively correlated true plots, but it does not improve the weak performance in negative correlation.
Figures 10 and 11 give a breakdown of the interactions between the response variables: plot detection, time and confidence. Figure 8 is the histogram of the time taken to record a response, split by plot design and sign of correlation and true plot detection. The time data was highly positively skewed with a maximum of 1760 seconds so the variable was log-transformed. We found that the subject could identify the true plot out of the overlaid line graph 33.6\% of the time and it took on average $e^{3.442}$ = 31.2 seconds to make a choice. Successful detections were slightly faster than unsuccessful decision, taking on average 28.7 seconds to pick the true plot and 32.6 seconds if a null was selected. The scatter plots performed much better, with the true plot picked 52.9\% of the time, with an average of 22.7 seconds required to make a decision. If the true plot was detected it took only 18.6 seconds, and 29.3 seconds for a null plot selection.

@@

Next we analysed the individual ability of each of our subjects. Figure 12 plots the accuracy of each subject on the y axis versus the average confidence on the x axis. While we expected that subjects with greater visual skills would report lower values of confidence, there does not appear to be any relationship between the two. The accuracy is normally distributed (Jarque-Bera p value = 0.418), with a mean of 42.5\% and a standard deviation of 17.8\%. The strongest subject successfully identified the true plot 91\% of the time, whilst the weakest was never able to identify the true plot. As all subjects were shown a balanced mix of true plot correlations, the extreme difference in individual visual ability is striking. Visual ability will be controlled via a subject specific random effect in the models in the next section.

@@

There are two major reasons why a subject may fail to detect the true plot out of a lineup. Either the plot design was unable to display correlation for that particular combination of factors (and the analyst would commit a Type II error by not rejecting the null hypothesis), or a null plot that had a greater correlation than the true plot was selected instead (and the analyst would commit a Type I error by rejecting the null hypothesis). We next can analyse these kinds of errors with the pin plots found in Figure 13. The frequency of selection of each plot in each lineup, either true or null, is represented by a pin, placed at the value of the correlation between the two variables shown in that plot. The correlation of the selected plot is on the x-axis and the frequency of selection is plotted on the y axis. The correlation of the simulated (true) data (r=-0.9, -0.7, -0.3, 0.3, 0.7, 0.9) in the plot, the sample size (t=12, 24, 48, 96), and the plot design (unsmoothed/smoothed, scatterplot/lineplot) is used to facet the plots. The blue pins indicate the true plots and the red pins indicate the null plots. If the plot design has the power to display correlation, we would expect the highest correlation plot to be selected, leading to a Type I error where a plot from the null distribution spuriously had high correlation. We find that this is generally not the case. The selections appear to have a wide range of correlations, indicating that if a subject selected a null plot it was typically not due to excess spurious correlation. Instead the failure to select the highest correlated plot indicates that a Type II error was committed rather than Type I as the plot designs did not have the power to differentiate between medium and low correlation plots. However, for the scatter plot, whenever there was a plot with a very strong correlation, usually in the form of the true plot, it was overwhelmingly selected. For the overlaid lines this mostly occurred only for plots with strong positive correlation, which is consistent with its general lack of power to reveal negative relationships in data.

\begin{figure}[htbp]
  \centering
<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
# correl <- read.csv("correlations.csv") # You have already read this
pin <- select(full, pic_id, response_0, actual_location, 
              test_param, smoothed, n) #full[,c(5, 6, 10, 15, 16, 17)]
# Need to use the tidyr commands instead of cSplit
# Here's where we will need to make some changes to incorporate the multiple picks
pin <- cSplit(pin, "response_0", direction="long")
counts <- data.frame(table(pin$pic_id, pin$response_0))
counts$Var1 <- as.integer(counts$Var1)
counts$Var2 <- as.integer(counts$Var2)
colnames(counts) <- c("lineup", "plot", "count")
pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
pin$r <- factor(pin$r)
colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
pin$base <- 0
pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))
pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
pin <- mutate(pin, percent = count/attempt)
ggplot(data=subset(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")

@@

\begin{figure}[htbp]
  \centering
Insert anomaly lineup 1
  \label{anom1}
  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707. The true plot in position 8 was not detected in five evaluations. However, plot 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225. When asked for a reason of their selections, subjects answered that the lines were closest together.}
\end{figure}

\begin{figure}[htbp]
  \centering
Insert anomaly lineup 2
  \label{anom2}
  \caption{A lineup of unsmoothed scatter plots with a sample size of 24. The true plot is located in position 4 with a correlation of -0.491, and was selected by one out of eleven observers (p = 0.33, see formula \ref{eq:1}). The plot in position 6 was selected by nine of these eleven observers (p < 0.001) despite having a correlation of -0.037. This null plot shows a unique clustering effect that draws in the observer despite not having any linear relationship in the data, an example of the over-interpretation of randomness. Subjects most often cited the clustering pattern formed as the reason for their selection. The final selection was position 8, with a correlation of -0.226 (p = 0.33).}
\end{figure}
